---
title: "Coverage parsing"
author: "Ivan Trus"
date: "2020-10-19"
output:
  html_document: default
  pdf_document: default
---
Execution start time: *`r Sys.time()`*
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = TRUE,
  warning = TRUE
)
library(vroom)
library(reshape)
```

# Initialization of the script and data loading
```{r init, message=FALSE}
Filelist <- list.files(pattern = "*.txt")
CountGenomes <- length(Filelist)
Threshold <- 400

# Defining functions
ListGaps <- function(Gaps) {
  Start <- append(Gaps[1], Gaps[which(diff(Gaps) > 1) + 1])
  Start
  End <- append(Gaps[which(diff(Gaps) > 1)], Gaps[length(Gaps)])
  End
  GapsSummary <- data.frame(Start, End)
  colnames(GapsSummary) <- c("Start", "End")
  return(GapsSummary)
}

Coverage <- vroom(Filelist, id = Filelist, col_names = "")
Coverage <- Coverage[c(1, 3, 4)]
colnames(Coverage)[1] <- "X1"
Coverage <- t(cast(Coverage, X1 ~ X2, value.var = "X3"))
colnames(Coverage) <- gsub(".txt", "", Filelist)
GenomeLength <- length(Coverage[, 1])

```

* Current working directory: **`r getwd()`**
* List of files to process: **`r Filelist`**
* Total number of files: **`r CountGenomes`**
* Threshold for coverage counting: **`r Threshold`** x
* Genome length: **`r GenomeLength`** nt

# Diagnostic data, chart and list of gaps for each sample
```{r main}
Results <- matrix(
  nrow = CountGenomes, ncol = 3,
  dimnames = list(1:CountGenomes, c(
    "Sample", "Read depth, x", "Coverage, %"))
)
Results <- as.data.frame(Results)
Results[, 1] <- colnames(Coverage)
AllGaps <- NA

summary(Coverage)

for (i in 1:CountGenomes) {
  Results[i, 2] <- round(mean(Coverage[, i]), digits = 4)
  Results[i, 3] <- round(100 - sum(Coverage[, i] < Threshold) / GenomeLength * 100, digits = 4)

  # We need to replace 0 with 1 to make it possible to be plotted fully on log scale
  GapsForLogChart <- Coverage[, i]
  GapsForLogChart[GapsForLogChart == 0] <- 1

  if(max(GapsForLogChart) < Threshold) YaxisLimit <- Threshold else YaxisLimit <- max(GapsForLogChart)
  plot(log10(GapsForLogChart),
    type = "l", xlab = "Nucleotide position in the genome",
    ylab = "Read depth (log10)", ylim = c(0, log10(YaxisLimit) * 1.05))
  text(x = 0, y = 1.05 * log10(YaxisLimit), colnames(Coverage)[i], pos = 4, font = 2, cex = 1.2)
  abline(h = log10(Threshold), col = "red", lwd = 3)
  GapsPositions <- as.numeric(names(subset(Coverage[, i], Coverage[, i] < Threshold)))
  AllGaps <- sort(c(AllGaps, setdiff(GapsPositions, AllGaps)))
  print(paste("The list of gaps in", colnames(Coverage)[i], "with the depth less than", Threshold, "x"))
  print(ListGaps(GapsPositions))
}
```

# Summary on read depth and coverage
```{r final}
Results
print(paste("Combined gaps with coverage depth less than", Threshold, "x."))
print(ListGaps(AllGaps))
print(paste("Intersection of coverage in all provided files:", round(100 - length(AllGaps) / GenomeLength * 100, digits = 4), "%"))
```

Execution end time: *`r Sys.time()`*
